---
title: "Imputation of missing covariates"
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
    toc_depth: 3
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br/>

[__Mark Scheuerell__](https://faculty.washington.edu/scheuerl/)  
_Fish Ecology Division, Northwest Fisheries Science Center, National Marine Fisheries Service, National Oceanic and Atmospheric Administration, Seattle, WA, USA_

__Karl Veggerby__  
_Ocean Associates, Seattle, WA, USA_

***

This is version `r paste0('0.',format(Sys.time(), '%y.%m.%d'))`.

***

# Background


# Requirements

```{r load_pkgs, message=FALSE}
## for analysis
library(MARSS)
## for plotting
library(viridisLite)
## for dir mgmt
library(here)
datadir <- here("data")
```

# Data munging

We begin by loading the data file with the temperature summaries by time (rows) and location (cols).

```{r load_covars}
## load obs covariates
cobs <- read.csv(file.path(datadir, "daily_mean_temp.csv"),
                 stringsAsFactors = FALSE)
## inspect data
head(cobs)
```

Let's simplify some names in the file and round the observations to the nearest 0.01.

```{r change_colnames}
## simplify colnames
colnames(cobs) <- gsub("data_source", replacement = "s", x = colnames(cobs))
colnames(cobs) <- gsub("daily_mean", replacement = "T", x = colnames(cobs))
## site abbrevs
sites <- gsub("T_", replacement = "", x = colnames(cobs)[10:16])
## simplify sources
cobs[cobs=="team carcass"] <- "car"
cobs[cobs=="achord_gordy"] <- "tag"
## round temps
cobs[,grep("T_", colnames(cobs))] <- round(cobs[,grep("T_", colnames(cobs))], 2)
```

The temperature data come from two different sources:

  1. the food web group (Sanderson et al.), and  
  2. the juvenile tagging group (Achord, Axel et al.).  
  
We need to split the data out by those groups.

```{r split_obs_grps}
## empty data frames
car <- tag <- matrix(NA, nrow(cobs), length(sites),
                     dimnames = list(NULL, sites))
## measurements
vals <- cobs[,-(1:9)]
## indices of data type
i_car <- cobs[,3:9] == "car"
i_tag <- cobs[,3:9] == "tag"
## group-specific data
car[i_car] <- vals[i_car]
tag[i_tag] <- vals[i_tag]
## drop any sites with all NA's
car <- car[,apply(car, 2, function(x) !all(is.na(x)))]
colnames(car) <- paste0(colnames(car), "_car")
tag <- tag[,apply(tag, 2, function(x) !all(is.na(x)))]
colnames(tag) <- paste0(colnames(tag), "_tag")
## regroup by sites & source
cobs_m <- cbind(car, tag)
cobs_m <- cobs_m[,sort(colnames(cobs_m))]
```

## Plot the data

```{r data_plot}
par(mai=c(0.9,0.9,0.1,0.1), omi=c(0,0,0,0))
## plot the data
matplot(cobs_m, type = "l", lty = "solid",
        lwd = 2, las = 1,
        ylab = "Temperature (C)", xaxt="n",
        col = plasma(ncol(cobs_m), alpha = 0.5, begin = 0, end = 1))
years <- unique(cobs$year)
n_yrs <- length(years)
n_days <- nrow(cobs) / n_yrs
axis(1,
     at = seq(n_days / 2, by = n_days, length.out = n_yrs),
     labels = years,
     tick = FALSE, line = -1,
     cex.axis = 0.7)
```

